---
title: "netCDF LOCA Data Import Data Block"
output: html_notebook
---

#1 Intro

This script is the reading section of pulling LOCA data.

## 1a. Loading Libraries


```{r}

  # Libraries 

  require("ncdf4")          # netCDF file access
  require("ncdf4.helpers")  # netCDF support files
  require("ggplot2")        # ggplot2 package (what could go wrong?)
  require("lubridate")      # data control
  require("reshape2")       # reshaping arrays
  require("abind")          # for concatenating data frames

```

# 2. Cresting Scenario Data

## 2a. Target Location, Latitude, and Longitude

```{r}

  # Select and Identify Targets.
  
  target_lon = -103.0573 + 360.   # degrees east 
  target_lat =   44.0435          # degrees north
  target_name = "Rapid City, SD"  # human station name
  target_abbr = "KRAP"            # simple character code (I use 4-letter ICAOs)

```

## 2b. Ensemble Members, Variables, and Scenarios

```{r}

  # Ensemble Members

  ensembles = c( "ACCESS1-0_r1i1p1",
                 "ACCESS1-3_r1i1p1",
                 "CCSM4_r6i1p1",
                 "CESM1-BGC_r1i1p1",
                 "CESM1-CAM5_r1i1p1",
                 "CMCC-CMS_r1i1p1",
                 "CMCC-CM_r1i1p1",
                 "CNRM-CM5_r1i1p1",
                 "CSIRO-Mk3-6-0_r1i1p1",
                 "CanESM2_r1i1p1",
                 "FGOALS-g2_r1i1p1",
                 "GFDL-CM3_r1i1p1",
                 "GFDL-ESM2G_r1i1p1",
                 "GFDL-ESM2M_r1i1p1",
                 "HadGEM2-AO_r1i1p1",
                 "HadGEM2-CC_r1i1p1",
                 "HadGEM2-ES_r1i1p1",
                 "IPSL-CM5A-LR_r1i1p1",
                 "IPSL-CM5A-MR_r1i1p1",
                 "MIROC-ESM-CHEM_r1i1p1",
                 "MIROC-ESM_r1i1p1",
                 "MIROC5_r1i1p1",
                 "MPI-ESM-LR_r1i1p1",
                 "MPI-ESM-MR_r1i1p1",
                 "MRI-CGCM3_r1i1p1",
                 "NorESM1-M_r1i1p1",
                 "bcc-csm1-1-m_r1i1p1" )

variables = c("pr", "tasmax", "tasmin")

scenarios = c("historical", "rcp45", "rcp85" )

```


# 2c. Opening OPeNDAP Remote File Information


```{r}

  # OPeNDAP Data Set to pull coordinates 

  root_url     = "http://kyrill.ias.sdsmt.edu:8080/thredds/dodsC/LOCA_SODAK"
  dataset_name = "SODAK_LOCA"


```


# 3. Pulling Metadata (Coordinates for time and location)

## 3a. Crack one History file and one RCP file open for reading

```{r}  
  
  # File Control and Opening
  
  var  = 1 # precip
  ens  = 1 # ensembles scenarios
  scen = 2 # RCP 4.5
  
  variable_rcp45  = paste(variables[var],  "_",
                          ensembles[ens],  "_",
                          scenarios[scen],
                          sep = "")
  
  thredd_URL_rcp45 = paste(root_url,        "/",
                           scenarios[scen], "/",
                           variables[var],  "/",
                           dataset_name,    "_",
                           variable_rcp45,  ".nc",
                           sep = "")

  scen = 1

  variable_histo  = paste(variables[var],   "_",
                          ensembles[ens],   "_",
                          scenarios[scen], 
                          sep = "")
  
  thredd_URL_histo = paste(root_url,        "/",
                           scenarios[scen], "/",
                           variables[var],  "/",
                           dataset_name,    "_",
                           variable_histo,  ".nc",
                           sep = "")


  # here we will open up the files
  #  we will try to include details in the argument list 
  #     for the functions we use in this demo

  nc.histo = nc_open(filename = thredd_URL_histo)  # name of the file
  nc.rcp45 = nc_open(filename = thredd_URL_rcp45)  # name of the file
  
```

## 3b. Get Spatial Coordiantes

```{r}

  # Extracting Latitude and Longitude Arrays
  
  lon =  ncvar_get(nc    = nc.rcp45, # netCDF file handle
                   varid =    "lon") # netCDF variable name
  
  lat =  ncvar_get(nc    = nc.rcp45, # netCDF file handle
                   varid =    "lat") # netCDF variable name

  # Get Point Coordinate Closest to Target Lat/Lon
  i_lon = which.min( abs(lon-target_lon) ) # which.min returns the index of the lowest value
  j_lat = which.min( abs(lat-target_lat) )
  
```

## 3c. Get Time Coordinates (need two sets, one for historical era, one for future era)

```{r}

# Extracting Time Coordinate Arrays

  # Historical Case
  
  time_histo = nc.get.time.series(f                            = nc.histo, # netCDF file handle
                                  time.dim.name                =   "time", # netCDF time coordinate name
                                  correct.for.gregorian.julian =    FALSE) # fix for julian calendar?

  # Convert from climate-science friendly PCICt format to R's prefered POSIXct format

  time_histo = as.POSIXct(time_histo)
  
  nt_histo   = length(time_histo)
  
  
  # Historical Case
  
  time_rcpXX = nc.get.time.series(f                            = nc.rcp45, # netCDF file handle
                                  time.dim.name                =   "time", # netCDF time coordinate name
                                  correct.for.gregorian.julian =    FALSE) # fix for julian calendar?

  # Convert from climate-science friendly PCICt format to R's prefered POSIXct format

  time_rcpXX = as.POSIXct(time_rcpXX)
  
  nt_rcpXX   = length(time_rcpXX)

  # we're closed with our first read so let's close thise files
  
  nc_close(nc = nc.rcp45)
  nc_close(nc = nc.histo)
  
```

## 3d. Get rid of unnecessary variables.

```{r}

# and clean things up... 
  
  remove(nc.rcp45)
  remove(nc.histo)
  remove(thredd_URL_rcp45)
  remove(thredd_URL_histo)
  remove(lon)
  remove(lat)
  remove(variable_histo)
  remove(variable_rcp45)

```

# 4.  Data Extraction of Time Series
  
Now for the Scary Part.  We are going to do a set of loops to pull in the data.  We'll need to do all of this in one {r} code block.

```{r}
# Data Import

  # Part 1:  Extract the LOCA Data into Columnmar Data Frames

  # For each Variable and Scenario, do the following
  #    Create Data Frame (time only for the given scenario)
  #    Loop through all files and extract the time series for our target point

```

# 4a. Historical Cases

```{r}

#### Historical Case 

    print("Processing Pre-2006 Scenarios")


    scen = 1 # historical
    ens  = 1 # ensemble member (for testing, this get looped)

    histo_pr = data.frame(Time = time_histo)
    
    histo_pr$Scenario = "Historical"
    
    histo_tx = histo_pr
    histo_tn = histo_pr

    histo_pr$Variable = "pr"
    histo_tx$Variable = "tasmax"
    histo_tn$Variable = "tasmin"

    
    for ( ens in 1:length(ensembles) )  {

      #####################################
      #
      # Precipitation
      #
      
      var  = 1 # precip (for testing, this get perscribed in loop)

      variable_name  = paste(variables[var],   "_",
                             ensembles[ens],   "_",
                             scenarios[scen], 
                             sep = "")

      print(paste("---processing    ", 
                  variable_name,
                  sep = " ")  ) 
  
      thredd_URL      = paste(root_url,        "/",
                              scenarios[scen], "/",
                              variables[var],  "/",
                              dataset_name,    "_",
                              variable_name,   ".nc",
                              sep = "")
      
      nc.file = nc_open(filename = thredd_URL)  # name of the file

      in_vector = ncvar_get(nc   =                     nc.file, # netCDF file handle
                            varid =              variable_name, # netCDF variable name
                            start = c(i_lon, j_lat,         1), # netCDF start indicies
                            count = c(     1,    1,  nt_histo)) # netCDF imported lengths
      
      histo_pr$in_vector = as.numeric(in_vector)
      
      names(histo_pr)[3+ens] = ensembles[ens]

      #      
      #####################################

      #####################################
      #
      # Max Air Temperature
      #
      
      var  = 2 # tasmax (for testing, this get perscribed in loop)

      variable_name  = paste(variables[var],   "_",
                             ensembles[ens],   "_",
                             scenarios[scen], 
                             sep = "")
      
      print(paste("---processing", 
                  variable_name,
                  sep = " ")  ) 
  
      thredd_URL      = paste(root_url,        "/",
                              scenarios[scen], "/",
                              variables[var],  "/",
                              dataset_name,    "_",
                              variable_name,   ".nc",
                              sep = "")
      
      nc.file = nc_open(filename = thredd_URL)  # name of the file

      in_vector = ncvar_get(nc   =                     nc.file, # netCDF file handle
                            varid =              variable_name, # netCDF variable name
                            start = c(i_lon, j_lat,         1), # netCDF start indicies
                            count = c(     1,    1,  nt_histo)) # netCDF imported lengths
      
      nc_close(nc.file)
      
      histo_tx$in_vector = as.numeric(in_vector)
      
      names(histo_tx)[3+ens] = ensembles[ens]

      #      
      #####################################
      
      #####################################
      #
      # Min Air Temperature
      #
      
      var  = 3 # tasmin (for testing, this get perscribed in loop)

      variable_name  = paste(variables[var],   "_",
                             ensembles[ens],   "_",
                             scenarios[scen], 
                             sep = "")
      
      print(paste("---processing", 
                  variable_name,
                  sep = " ")  ) 
  
      thredd_URL      = paste(root_url,        "/",
                              scenarios[scen], "/",
                              variables[var],  "/",
                              dataset_name,    "_",
                              variable_name,   ".nc",
                              sep = "")
      
      nc.file = nc_open(filename = thredd_URL)  # name of the file

      in_vector = ncvar_get(nc   =                     nc.file, # netCDF file handle
                            varid =              variable_name, # netCDF variable name
                            start = c(i_lon, j_lat,         1), # netCDF start indicies
                            count = c(     1,    1,  nt_histo)) # netCDF imported lengths
      
      nc_close(nc.file)

      histo_tn$in_vector = as.numeric(in_vector)
      
      names(histo_tn)[3+ens] = ensembles[ens]

      #      
      #####################################
      
    }
    
    print(" ")
    
    # cleanup
    
    remove(in_vector)
    remove(variable_name)
    remove(thredd_URL)
    remove(var)
    remove(scen)
    remove(ens)

```

# 4b. RCP 4.5 Cases

```{r}

#### RCP 4.5  Case 

    print("Processing RCP 4.5 CMIP-5 Scenario")

    scen = 2 # RCP 4.5

    rcp45_pr = data.frame(Time = time_rcpXX)
    
    rcp45_pr$Scenario = "RCP 4.5"
    
    rcp45_tx = rcp45_pr
    rcp45_tn = rcp45_pr

    rcp45_pr$Variable = "pr"
    rcp45_tx$Variable = "tasmax"
    rcp45_tn$Variable = "tasmin"

    for ( ens in 1:length(ensembles) )  {

      #####################################
      #
      # Precipitation
      #
      
      var  = 1 # precip (for testing, this get perscribed in loop)

      variable_name  = paste(variables[var],   "_",
                             ensembles[ens],   "_",
                             scenarios[scen], 
                             sep = "")
      
      print(paste("---processing    ", 
                  variable_name,
                  sep = " ")  ) 
  
      thredd_URL      = paste(root_url,        "/",
                              scenarios[scen], "/",
                              variables[var],  "/",
                              dataset_name,    "_",
                              variable_name,   ".nc",
                              sep = "")
      
      nc.file = nc_open(filename = thredd_URL)  # name of the file

      in_vector = ncvar_get(nc   =                     nc.file, # netCDF file handle
                            varid =              variable_name, # netCDF variable name
                            start = c(i_lon, j_lat,         1), # netCDF start indicies
                            count = c(     1,    1,  nt_rcpXX)) # netCDF imported lengths
      
      rcp45_pr$in_vector = as.numeric(in_vector)
      
      names(rcp45_pr)[3+ens] = ensembles[ens]

      #      
      #####################################

      #####################################
      #
      # Max Air Temperature
      #
      
      var  = 2 # tasmax (for testing, this get perscribed in loop)

      variable_name  = paste(variables[var],   "_",
                             ensembles[ens],   "_",
                             scenarios[scen], 
                             sep = "")
      
      print(paste("---processing", 
                  variable_name,
                  sep = " ")  ) 
  
      thredd_URL      = paste(root_url,        "/",
                              scenarios[scen], "/",
                              variables[var],  "/",
                              dataset_name,    "_",
                              variable_name,   ".nc",
                              sep = "")
      
      nc.file = nc_open(filename = thredd_URL)  # name of the file

      in_vector = ncvar_get(nc   =                     nc.file, # netCDF file handle
                            varid =              variable_name, # netCDF variable name
                            start = c(i_lon, j_lat,         1), # netCDF start indicies
                            count = c(     1,    1,  nt_rcpXX)) # netCDF imported lengths
      
      nc_close(nc.file)
      
      rcp45_tx$in_vector = as.numeric(in_vector)
      
      names(rcp45_tx)[3+ens] = ensembles[ens]

      #      
      #####################################
      
      #####################################
      #
      # Min Air Temperature
      #
      
      var  = 3 # tasmin (for testing, this get perscribed in loop)

      variable_name  = paste(variables[var],   "_",
                             ensembles[ens],   "_",
                             scenarios[scen], 
                             sep = "")
      
      print(paste("---processing", 
                  variable_name,
                  sep = " ")  ) 
  
      thredd_URL      = paste(root_url,        "/",
                              scenarios[scen], "/",
                              variables[var],  "/",
                              dataset_name,    "_",
                              variable_name,   ".nc",
                              sep = "")
      
      nc.file = nc_open(filename = thredd_URL)  # name of the file

      in_vector = ncvar_get(nc   =                     nc.file, # netCDF file handle
                            varid =              variable_name, # netCDF variable name
                            start = c(i_lon, j_lat,         1), # netCDF start indicies
                            count = c(     1,    1,  nt_rcpXX)) # netCDF imported lengths
      
      nc_close(nc.file)

      rcp45_tn$in_vector = as.numeric(in_vector)
      
      names(rcp45_tn)[3+ens] = ensembles[ens]

      #      
      #####################################
      
    }
    
    print(" ")

    # cleanup
    
    remove(in_vector)
    remove(variable_name)
    remove(thredd_URL)
    remove(var)
    remove(scen)
    remove(ens)
    
```


# 4c. RCP 8.5 Cases

```{r}

#### RCP 8.5  Case 

    print("Processing RCP 8.5 CMIP-5 Scenario")

    scen = 3 # RCP 8.5

    rcp85_pr = data.frame(Time = time_rcpXX)
    
    rcp85_pr$Scenario = "RCP 8.5"
  
    rcp85_tx = rcp85_pr
    rcp85_tn = rcp85_pr

    rcp85_pr$Variable = "pr"
    rcp85_tx$Variable = "tasmax"
    rcp85_tn$Variable = "tasmin"

    for ( ens in 1:length(ensembles) )  {

      #####################################
      #
      # Precipitation
      #
      
      var  = 1 # precip (for testing, this get perscribed in loop)

      variable_name  = paste(variables[var],   "_",
                             ensembles[ens],   "_",
                             scenarios[scen], 
                             sep = "")
      
      print(paste("---processing    ", 
                  variable_name,
                  sep = " ")  ) 
  
      thredd_URL      = paste(root_url,        "/",
                              scenarios[scen], "/",
                              variables[var],  "/",
                              dataset_name,    "_",
                              variable_name,   ".nc",
                              sep = "")
      
      nc.file = nc_open(filename = thredd_URL)  # name of the file

      in_vector = ncvar_get(nc   =                     nc.file, # netCDF file handle
                            varid =              variable_name, # netCDF variable name
                            start = c(i_lon, j_lat,         1), # netCDF start indicies
                            count = c(     1,    1,  nt_rcpXX)) # netCDF imported lengths
      
      rcp85_pr$in_vector = as.numeric(in_vector)
      
      names(rcp85_pr)[3+ens] = ensembles[ens]

      #      
      #####################################
            
      #####################################
      #
      # Max Air Temperature
      #
      
      var  = 2 # tasmax (for testing, this get perscribed in loop)

      variable_name  = paste(variables[var],   "_",
                             ensembles[ens],   "_",
                             scenarios[scen], 
                             sep = "")
      
      print(paste("---processing", 
                  variable_name,
                  sep = " ")  ) 
  
      thredd_URL      = paste(root_url,        "/",
                              scenarios[scen], "/",
                              variables[var],  "/",
                              dataset_name,    "_",
                              variable_name,   ".nc",
                              sep = "")
      
      nc.file = nc_open(filename = thredd_URL)  # name of the file

      in_vector = ncvar_get(nc   =                     nc.file, # netCDF file handle
                            varid =              variable_name, # netCDF variable name
                            start = c(i_lon, j_lat,         1), # netCDF start indicies
                            count = c(     1,    1,  nt_rcpXX)) # netCDF imported lengths
      
      nc_close(nc.file)
      
      rcp85_tx$in_vector = as.numeric(in_vector)
      
      names(rcp85_tx)[3+ens] = ensembles[ens]

      #      
      #####################################
            
      #####################################
      #
      # Min Air Temperature
      #
      
      var  = 3 # tasmin (for testing, this get perscribed in loop)

      variable_name  = paste(variables[var],   "_",
                             ensembles[ens],   "_",
                             scenarios[scen], 
                             sep = "")
      
      print(paste("---processing", 
                  variable_name,
                  sep = " ")  ) 
  
      thredd_URL      = paste(root_url,        "/",
                              scenarios[scen], "/",
                              variables[var],  "/",
                              dataset_name,    "_",
                              variable_name,   ".nc",
                              sep = "")
      
      nc.file = nc_open(filename = thredd_URL)  # name of the file

      in_vector = ncvar_get(nc   =                     nc.file, # netCDF file handle
                            varid =              variable_name, # netCDF variable name
                            start = c(i_lon, j_lat,         1), # netCDF start indicies
                            count = c(     1,    1,  nt_rcpXX)) # netCDF imported lengths
      
      nc_close(nc.file)

      rcp85_tn$in_vector = as.numeric(in_vector)
      
      names(rcp85_tn)[3+ens] = ensembles[ens]

      #      
      #####################################
      
    }
    
    print(" ")
        
    # cleanup
    
    remove(in_vector)
    remove(variable_name)
    remove(thredd_URL)
    remove(var)
    remove(scen)
    remove(ens)
    
```

```{r}

save(dataset_name,
     target_name,
     target_abbr,
     target_lon,
     target_lat,
     histo_pr,
     histo_tx,
     histo_tn,
     rcp45_pr,
     rcp45_tx,
     rcp45_tn,
     rcp85_pr,
     rcp85_tx,
     rcp85_tn,
     file = paste("LOCA_1_",
                  target_abbr,
                  ".Rdata",
                  sep = ""))

```


# 5.  Creating Data Frames

```{r}

# Data Frame Creation

  # We have arrays for each scenario.  We will now merge them into a single data frame.

```

## 5a. Merging Historical Cases

```{r}

  # We start with the history case with the melt function

  histo.pr =   melt(data          = histo_pr,        # your array
                    na.rm         = FALSE,           # use missing values
                    id.vars       =  c(    "Time",   # names of your two dimensions    
                                       "Variable",
                                       "Scenario"),  
                    variable.name = "Ensemble",      # new for new dimensions
                    value.name    = "Daily_Precip")  # the final name of your aray value
  
  histo.tx =   melt(data          = histo_tx,         # your array
                    na.rm         = FALSE,            # use missing values
                    id.vars       =  c(    "Time",    # names of your two dimensions    
                                       "Variable",
                                       "Scenario"),  
                    variable.name = "Ensemble",       # new for new dimensions
                    value.name    = "Daily_Max_Temp") # the final name of your aray value

  histo.tn =   melt(data          = histo_tn,         # your array
                    na.rm         = FALSE,            # use missing values
                    id.vars       =  c(    "Time",    # names of your two dimensions    
                                       "Variable",
                                       "Scenario"),  
                    variable.name = "Ensemble",       # new for new dimensions
                    value.name    = "Daily_Min_Temp") # the final name of your aray value
  
  # Now we put them all into one data frame for all historical period data
  
  historical = histo.pr
  
  historical$Daily_Max_Temp = histo.tx$Daily_Max_Temp
  historical$Daily_Min_Temp = histo.tn$Daily_Min_Temp

  # And Clean the Mess

  remove(histo.pr)
  remove(histo.tx)
  remove(histo.tn)

```

## 5b. Merging RCP 4.5 Cases

```{r}

  # Now we process the RCP 4.5 projection case with the melt function

  rcp45.pr =   melt(data          = rcp45_pr,        # your array
                    na.rm         = FALSE,           # use missing values
                    id.vars       =  c(    "Time",   # names of your two dimensions    
                                       "Variable",
                                       "Scenario"),  
                    variable.name = "Ensemble",      # new for new dimensions
                    value.name    = "Daily_Precip")  # the final name of your aray value

  rcp45.tx =   melt(data          = rcp45_tx,         # your array
                    na.rm         = FALSE,            # use missing values
                    id.vars       =  c(    "Time",    # names of your two dimensions    
                                       "Variable",
                                       "Scenario"),  
                    variable.name = "Ensemble",       # new for new dimensions
                    value.name    = "Daily_Max_Temp") # the final name of your aray value

  rcp45.tn =   melt(data          = rcp45_tn,         # your array
                    na.rm         = FALSE,            # use missing values
                    id.vars       =  c(   "Time",    # names of your two dimensions    
                                       "Variable",
                                       "Scenario"),  
                    variable.name = "Ensemble",       # new for new dimensions
                    value.name    = "Daily_Min_Temp") # the final name of your aray value
  
  # Now we put them all into one data frame for all RCP 4.5 projection data
  
  rcp45 = rcp45.pr
  
  rcp45$Daily_Max_Temp = rcp45.tx$Daily_Max_Temp
  rcp45$Daily_Min_Temp = rcp45.tn$Daily_Min_Temp

  # And Clean the Mess

  remove(rcp45.pr)
  remove(rcp45.tx)
  remove(rcp45.tn)

```

## 5c. Merging RCP 8.5 Cases
 
```{r}

  # Finally we process the RCP 8.5 projection case with the melt function

  rcp85.pr =   melt(data          = rcp85_pr,        # your array
                    na.rm         = FALSE,           # use missing values
                    id.vars       =  c(    "Time",   # names of your two dimensions    
                                       "Variable",
                                       "Scenario"),  
                    variable.name = "Ensemble",      # new for new dimensions
                    value.name    = "Daily_Precip")  # the final name of your aray value

  rcp85.tx =   melt(data          = rcp85_tx,         # your array
                    na.rm         = FALSE,            # use missing values
                    id.vars       =  c(    "Time",    # names of your two dimensions    
                                       "Variable",
                                       "Scenario"),  
                    variable.name = "Ensemble",       # new for new dimensions
                    value.name    = "Daily_Max_Temp") # the final name of your aray value

  rcp85.tn =   melt(data          = rcp85_tn,         # your array
                    na.rm         = FALSE,            # use missing values
                    id.vars       =  c(    "Time",    # names of your two dimensions    
                                       "Variable",
                                       "Scenario"),  
                    variable.name = "Ensemble",       # new for new dimensions
                    value.name    = "Daily_Min_Temp") # the final name of your aray value
  
  # Now we put them all into one data frame for all RCP 8.5 projection data
  
  rcp85 = rcp85.pr
  
  rcp85$Daily_Max_Temp = rcp85.tx$Daily_Max_Temp
  rcp85$Daily_Min_Temp = rcp85.tn$Daily_Min_Temp

  # And Clean the Mess

  remove(rcp85.pr)
  remove(rcp85.tx)
  remove(rcp85.tn)

```

```{r}
save(dataset_name,
     target_name,
     target_abbr,
     target_lon,
     target_lat,
     historical,
     rcp45,
     rcp85,
     file = paste("LOCA_2_",
                  target_abbr,
                  ".Rdata",
                  sep = ""))

```


## 5d. Concatenate (bind) ALL of the cases into a single dataframe

```{r}

 # to merge we use the rbind() function

loca = rbind(historical,
             rcp45,
             rcp85)

remove(historical)
remove(rcp45)
remove(rcp85)

```

## 5e. Save File For Now

```{r}

save(dataset_name,
     target_name,
     target_abbr,
     target_lon,
     target_lat,
     loca, 
     file = paste("LOCA_",
                  target_abbr,
                  ".Rdata",
                  sep = ""))

```
